---
title: "BBets"
author: "Angelica Romero"  
output: html_document
date: "2024-09-27"     
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, (month, day).        
a)Find the probability that at least two student have the same birthday.     
b)If 26 students in the class take the bet how much is Dr. J expected to win/lose.     
c)How much should Dr. J wager against the student's $1 to make this a fair bet?         

2. Assume class size is 27 instead of 35.  
a) With the same bet ($1 to $1) what is the probability Dr. J wins the bet now.  
b) If 20 of the 27 students make the bet, what is Dr. J's expected profit/loss?

3. What size class (how many students) makes the $1 againt $1 a fair bet?       


```{r setup, include=FALSE}
set.seed(123)  #Set seed for reproducibility
```

```{r}
# Function to simulate if two or more people share a birthday
same_birthday_prob <- function(class_size, n_sim) {
 count_same_birthday <- 0 
 for (i in 1:n_sim) { 
   birthdays <- sample(1:365, class_size, replace = TRUE) #Generate random birthdays
   if (length(unique(birthdays)) < class_size){ 
     count_same_birthday <- count_same_birthday + 1
 }
}
return(count_same_birthday / n_sim) # Return probability
} 

# Simulation settings
 class_size <- 35
 n_sim <- 1000
 n_trials <- 100
 
 # Vector to store the probabilities from each trial
 probs <- numeric(n_trials)
 
 # Run the simulation 100 times
 for (i in 1:n_trials) {
   probs[i] <- same_birthday_prob(class_size, n_sim)
 }
 
```


## mean and std dev

```{r, echo=F}
# Print mean, std of probabilities

mean_probs <- mean(probs) 
sd_probs <- sd(probs)

cat("Mean Probability:", mean_probs, "\n")
cat("Standard Deviation of Probabilities:", sd_probs, "\n")

```

The probability that Dr.J wins is .813. 

```{r}
EV= mean_probs*1*26
cat("expected profit =", EV)

correctbet = mean_probs/(1-mean_probs)
cat("correct bet=", correctbet)

```

```{r}
hist(probs, main = "Histogram of Probailities", xlab = "Probability", col = "lightblue", border = "black")

```

```{r}
 # Calculate 95% Confidence Interval
error_margin <- qt(0.975, df=n_trials-1) * sd_probs / sqrt(n_trials)
conf_interval <- c(mean_probs - error_margin, mean_probs + error_margin)

cat("95% Confidence Interval for the Mean Probability:[", conf_interval[1], conf_interval[2], "]\n")

```

The confidence interval is 0.811 to 0.816. 

Get answers for a smaller class

```{r}
# Simulation settings
class_size <- 27
n_sim <- 1000
n_trials <- 100

#Vector to store the probabilities from each trial 
probs <- numeric(n_trials) 

#Run the simulation 100 times 
for (i in 1:n_trials) {
  probs[i] <- same_birthday_prob(class_size, n_sim) 
}

cat("proability of win =" , mean(probs)) 
```
```{r}
EV=mean(probs)*1*20
cat("expected profit =" , EV) 
```

## mean and std dev

```{r echo=F}
#Simulation settings
class_size <- 27
n_sim <- 1000
n_trials <- 100

#Vector to store the probabilities from each trial
probs <- numeric(n_trials) 

#Run the simulation 100 times
for (i in 1:n_trials) {
  probs[i] <- same_birthday_prob(class_size, n_sim) 
} 

cat("probaility of win =" , mean(probs)) 

```




```{r}
# Simulation settings 
class_size <- 23
n_sim <- 1000
n_trials <- 100 

#Vector to store the probabilities from each trial
probs <- numeric(n_trials) 

#Run the simulation 100 times
for (i in 1:n_trials) {
  probs[i] <- same_birthday_prob(class_size, n_sim)
}

cat("probability of win =", mean(probs)) 

```
The number of fair students for a fair game 1:1 is 23

## Explanation: 

- **`same_birthday_prob` function**: Simulates whether at least two people in a group of class `class_size` share a birthday, repeated `n_sim` times. It returns the proportion of simulations where a birthday match occurs. 
- **`probs` vector**: Stores the results of running the birthday problem simulation 100 times.
- **Mean, standard deviation, and histogram**: Summarizes and visualizes the distribution of probabilities.
- **Confidence interval**: Calculates a 95% confidence interval for the mean of the simulated probabilities.
 
