---
title: "BBets"
author: "Angelica Romero"
output: html_document
date: "2024-09-27"
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, (month, day).        
a)Find the probability that at least two student have the same birthday.     
b)If 26 students in the class take the bet how much is Dr. J expected to win/lose.     
c)How much should Dr. J wager against the student's $1 to make this a fair bet?         

2. Assume class size is 27 instead of 35.  
a) With the same bet ($1 to $1) what is the probability Dr. J wins the bet now.  
b) If 20 of the 27 students make the bet, what is Dr. J's expected profit/loss?

3. What size class (how many students) makes the $1 againt $1 a fair bet?


```{r setup, include=FALSE}
set.seed(123)  #Set seed for reproducibility
```

```{r}
# Function to simulate if two or more people share a birthday
same_birthday_prob <- function(class_size, n_sim) {
  
}
```

